---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(sqldf)


here::here()

candy_clean <- read_csv(here("clean_data/candy_clean.csv"), guess_max = 9349)

# pivoting the data longer to have the data tidy for some of the analysis 

candy_clean_longer <- candy_clean %>% 
  pivot_longer(
    cols = ("100_grand_bar":"york_peppermint_patties"),
    names_to = "candy",
    values_to = "rating"
  )
```


Analysis questions

Q1 What is the total number of candy rating given across the three years. (number of candy rating, not number of raters. Don’t count missing values)

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(rating) %>% 
  filter(!is.na(rating)) %>% 
  summarise(total_rating = n())

## 2. wider table

### function to pick up all missing values per each column

nonNAs <- function(x) {
  as.vector(apply(x, 2, function(x) length(which(!is.na(x)))))
}

### subsetting data frame that contains only candies (column 7 onward)

candies <- candy_clean %>% 
  select(7:114)

### sum of all missing values 

sum(nonNAs(candies))
```


Q2 What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(age, going_out_trick_or_treating_yourself) %>% 
  filter(!is.na(going_out_trick_or_treating_yourself)) %>% 
  group_by(going_out_trick_or_treating_yourself) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))

## 2. wider table

candy_clean %>% 
  filter(going_out_trick_or_treating_yourself == "Yes" | going_out_trick_or_treating_yourself == "No") %>% 
  group_by(going_out_trick_or_treating_yourself) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))
```


Q3 For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(candy, rating) %>% 
  filter(grepl("bar", candy)) %>% 
  filter(!is.na(rating)) %>% 
  group_by(rating, candy) %>% 
  summarise(count = n()) %>% 
  filter(count == max(count))
  

## 2. wider table

# subsetting only candies that contain "bar" in their name
# note. n() gives the current group size
# pivoting into two columns, one with candy bar names, the second one with rating
# getting total of each rating type per each candy

candy_clean %>% 
  select(contains("bar")) %>% 
  pivot_longer(
    cols = 1:8,
    names_to = "candy_bars",
    values_to = "rating"
  ) %>% 
  filter(!is.na(rating)) %>% 
  group_by(rating, candy_bars) %>% 
  summarise(count = n()) %>% 
  filter(count == max(count))

# note can use function groups() to track what groups came through to a particular point
```

Q4 How many people rated Starburst as despair?

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(candy, rating) %>% 
  filter(candy == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(total = n())

## 2. wider table

candy_clean %>% 
  select(starburst) %>% 
  filter(starburst == "DESPAIR") %>% 
  summarise(total_starburst_despair = n())
```


# For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
## 1. longer table

candy_clean_counting_longer <- candy_clean_longer %>% 
  mutate(rating = case_when(
      rating == "JOY" ~ 1,
      rating == "DESPAIR" ~ -1,
      rating == "MEH" ~ 0,
    )
  )


## 2. wider table

### splitting data frame candy_clean into two data frames to be able to change rating to -1, 0, 1

candy_clean_sub_id <- candy_clean %>% 
  select(1:6)

candy_clean_counting_sub <- candy_clean %>% 
  select(7:ncol(candy_clean)) %>% 
  mutate_all(
    funs(case_when(
      . == "JOY" ~ 1,
      . == "DESPAIR" ~ -1,
      . == "MEH" ~ 0,
    ))) 

### and binding them back together

candy_clean_counting_wider <- cbind(candy_clean_sub_id, candy_clean_counting_sub) 
```


Q5 What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_clean_counting_longer %>% 
  select(candy, gender, rating) %>% 
  filter(!is.na(gender)) %>% 
  filter(!is.na(rating)) %>% 
  filter(grepl("bar", candy)) %>% 
  group_by(gender, candy) %>%  # summarise removes candy
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  slice_max(avg_rating) # slice_max respects group_by
```


Q6 What was the most popular candy bar in each year?

```{r}
candy_clean_counting_longer %>% 
  select(year, candy, rating) %>% 
  filter(grepl("bar", candy)) %>% 
  filter(!is.na(rating)) %>% 
  group_by(year, candy) %>% 
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  slice_max(avg_rating)
```


Q7 What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
candy_clean_counting_longer %>% 
  select(country, candy, rating) %>% 
  filter(grepl("bar", candy)) %>% 
  filter(!is.na(rating)) %>% 
  filter(!is.na(country)) %>% 
  mutate(
    country_group = case_when(
      country == "USA" ~ "USA",
      country == "UK"  ~ "UK",
      country == "Canada" ~ "Canada",
      TRUE ~ "others"
    )
  ) %>% 
  select(country_group, candy, rating) %>% 
  group_by(country_group, candy) %>% 
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  slice_max(avg_rating) 
```

