---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(sqldf)


here::here()

candy_clean <- read_csv(here("clean_data/candy_clean.csv"), guess_max = 9349)

glimpse(candy_clean) %>% head(10)
```


Analysis questions

Q1 What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
#JOY, DESPAIR or MEH

# function to pick up all missing values per each column

nonNAs <- function(x) {
  as.vector(apply(x, 2, function(x) length(which(!is.na(x)))))
}

# subsetting data frame that contains only candies (column 8 onward)

candies <- candy_clean %>% 
  select(7:117)

## sum of all missing values 

total_NA_candies <- sum(nonNAs(candies))
  
## number of all data points 

total_candy_ratings_incl_NA <- nrow(candies) * ncol(candies)

## total number of candy ratings given across the three years

total_ratings = total_candy_ratings_incl_NA - total_NA_candies
total_ratings
```


Q2 What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
candy_clean %>% 
  filter(going_out_trick_or_treating_yourself == "Yes" | going_out_trick_or_treating_yourself == "No") %>% 
  group_by(going_out_trick_or_treating_yourself) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))
```


Q3 For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
# subsetting only candies that contain "bar" in their name
# note. n() gives the current group size
# pivoting into two columns, one with candy bar names, the second one with rating
# getting total of each rating type per each candy

candy_bars <- candy_clean %>% 
  select(contains("bar")) %>% 
  pivot_longer(
    cols = 1:8,
    names_to = "candy_bars",
    values_to = "rating"
  ) %>% 
  group_by(candy_bars, rating) %>% 
  summarise(count = n())


candy_bars %>% 
  filter(rating == "JOY") %>% 
  arrange(desc(count)) %>% 
  head(1)

candy_bars %>% 
  filter(rating == "DESPAIR") %>% 
  arrange(desc(count)) %>% 
  head(1)

candy_bars %>% 
  filter(rating == "MEH") %>% 
  arrange(desc(count)) %>% 
  head(1)
```

Q4 How many people rated Starburst as despair?

```{r}
candy_clean %>% 
  select(starburst) %>% 
  filter(starburst == "DESPAIR") %>% 
  summarise(total_starburst_despair = n())
```


# For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
# splitting data frame candy_clean into two data frames to be able to change rating to -1, 0, 1

candy_clean_sub_id <- candy_clean %>% 
  select(1:6)

candy_clean_counting_sub <- candy_clean %>% 
  select(7:ncol(candy_clean)) %>% 
  mutate_all(
    funs(case_when(
      . == "JOY" ~ 1,
      . == "DESPAIR" ~ -1,
      . == "MEH" ~ 0,
    ))) 

# and binding them back together

candy_clean_counting <- cbind(candy_clean_sub_id, candy_clean_counting_sub) 
```


Q5 What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_clean_counting %>% 
  select(gender, contains("bar")) %>% 
  filter(!is.na(gender)) %>% 
  group_by(gender) %>% 
  summarise_all(~sum(., na.rm = TRUE))


    
```


Q6 What was the most popular candy bar in each year?

```{r}
candy_clean_counting %>% 
  select(year, contains("bar")) %>% 
  group_by(year) %>% 
  summarise_all(~sum(., na.rm = TRUE))




```


Q7 What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}

```

