---
title: "Dirty data project - Halloween Candy"
output: html_notebook
---

# Dirty data project 
## Task 4 - Halloween Candy

We are dealing with data obtained from three independent surveys.

These surveys were carried on in years 2015, 2016 and 2017.

Respondents from all over the world were ask to rank various candies based on their subjective opinion.

They were offered three options as an answer - *joy*, *despair*, or *meh*

There were also additional questions like what was the respondents' age, what was their country of origin, whether they prefer name Betty or Veronica etc. These questions had no default set of answers and it was up to the respondents to type in whatever they found appropriate.

Some candy types, as well as additional questions, did only appear in one or two surveys.

As a result, the raw data entering the analysis differed significantly and required some deep cleaning first.

More information on the data can be found 
[here](https://www.scq.ubc.ca/so-much-candy-data-seriously/).

### The project structure:

#### Data cleaning

* *data\_cleaning\_script\_task\_4.R* - combining the three data frames together and saving the cleaned data into a *candy\_clean.csv* file
 
#### Data analysis

* _candy\_analysis.Rmd_ - reading in the *candy\_clean.csv* file and carrying out an analysis

Both files are thoroughly commented on what steps have been taken.


```{r}
library(tidyverse)
library(here)
library(sqldf)
```

```{r}
here::here()

candy_clean <- read_csv(here("clean_data/candy_clean.csv"), guess_max = 9349)

# pivoting the data into longer format to have the data in "tidier" format for some of the analysis 

candy_clean_longer <- candy_clean %>% 
  pivot_longer(
    cols = ("100_grand_bar":"york_peppermint_patties"),
    names_to = "candy",
    values_to = "rating"
  )
```


## Analysis questions

### Q1 What is the total number of candy rating given across the three years. (number of candy rating, not number of raters. Don’t count missing values)

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(rating) %>% 
  filter(!is.na(rating)) %>% 
  summarise(total_rating = n())

## 2. wider table

### function to pick up all missing values per each column

nonNAs <- function(x) {
  as.vector(apply(x, 2, function(x) length(which(!is.na(x)))))
}

### subsetting data frame that contains only candies (column 7 onward)

candies <- candy_clean %>% 
  select(7:114)

### sum of all missing values 

sum(nonNAs(candies))
```


### Q2 What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(age, going_out_trick_or_treating_yourself) %>% 
  filter(!is.na(going_out_trick_or_treating_yourself)) %>% 
  group_by(going_out_trick_or_treating_yourself) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))

## 2. wider table

candy_clean %>% 
  filter(going_out_trick_or_treating_yourself == "Yes" | going_out_trick_or_treating_yourself == "No") %>% 
  group_by(going_out_trick_or_treating_yourself) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))
```


### Q3 For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
# subsetting only candies that contain "bar" in their name

## 1. longer table

candy_clean_longer %>% 
  select(candy, rating) %>% 
  filter(grepl("bar", candy)) %>% 
  filter(!is.na(rating)) %>% 
  group_by(rating, candy) %>% 
  summarise(count = n()) %>% 
  filter(count == max(count))
  
## 2. wider table

### pivoting into two columns, one with candy bar names, the second one with rating
### getting total of each rating type per each candy

candy_clean %>% 
  select(contains("bar")) %>% 
  pivot_longer(
    cols = 1:8,
    names_to = "candy_bars",
    values_to = "rating"
  ) %>% 
  filter(!is.na(rating)) %>% 
  group_by(rating, candy_bars) %>% 
  summarise(count = n()) %>% 
  filter(count == max(count))

#### note.include function groups() to track what groups came through to a particular point (handy especially after group_by step)
```

### Q4 How many people rated Starburst as despair?

```{r}
## 1. longer table

candy_clean_longer %>% 
  select(candy, rating) %>% 
  filter(candy == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(total = n())

## 2. wider table

candy_clean %>% 
  select(starburst) %>% 
  filter(starburst == "DESPAIR") %>% 
  summarise(total_starburst_despair = n())
```


### For the next three questions, count despair as -1, joy as +1 and meh as 0.

```{r}
## 1. longer table

candy_clean_counting_longer <- candy_clean_longer %>% 
  mutate(rating = case_when(
      rating == "JOY" ~ 1,
      rating == "DESPAIR" ~ -1,
      rating == "MEH" ~ 0,
    )
  )

## 2. wider table

### splitting data frame candy_clean into two data frames to be able to change rating to -1, 0, 1

candy_clean_sub_id <- candy_clean %>% 
  select(1:6)

candy_clean_counting_sub <- candy_clean %>% 
  select(7:ncol(candy_clean)) %>% 
  mutate_all(
    funs(case_when(
      . == "JOY" ~ 1,
      . == "DESPAIR" ~ -1,
      . == "MEH" ~ 0,
    ))) 

### and binding them back together

candy_clean_counting_wider <- cbind(candy_clean_sub_id, candy_clean_counting_sub) 
```


### Q5 What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_clean_counting_longer %>% 
  select(candy, gender, rating) %>% 
  filter(!is.na(gender)) %>% 
  filter(!is.na(rating)) %>% 
  filter(grepl("bar", candy)) %>% 
  group_by(gender, candy) %>%  # summarise removes candy
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  slice_max(avg_rating) # slice_max respects group_by
```

### Q6 What was the most popular candy bar in each year?

```{r}
candy_clean_counting_longer %>% 
  select(year, candy, rating) %>% 
  filter(grepl("bar", candy)) %>% 
  filter(!is.na(rating)) %>% 
  group_by(year, candy) %>% 
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  slice_max(avg_rating)
```

### Q7 What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?

```{r}
candy_clean_counting_longer %>% 
  select(country, candy, rating) %>% 
  filter(grepl("bar", candy)) %>% 
  filter(!is.na(rating)) %>% 
  filter(!is.na(country)) %>% 
  mutate(
    country_group = case_when(
      country == "USA" ~ "USA",
      country == "UK"  ~ "UK",
      country == "Canada" ~ "Canada",
      TRUE ~ "others"
    )
  ) %>% 
  select(country_group, candy, rating) %>% 
  group_by(country_group, candy) %>% 
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  slice_max(avg_rating) 
```

### Candy Hierarchy - all the candies sorted from the most popular to the least popular

#### an overall Halloween Candy winner accross the three years was any_full_sized_candy_bar, while a broken_glow_stick took the last place :)

```{r}
candy_clean_counting_longer %>% 
  select(candy, rating) %>% 
  filter(!is.na(rating)) %>% 
  group_by(candy) %>% 
  summarise(avg_rating = mean(rating, na.rm = TRUE)) %>% 
  arrange(desc(avg_rating))
```

